Revisiting the hydrodynamic modulation of short surface waves by longer waves

### 2025-02-15 paper.tex
Compute wave action conservation criteria
WIP wave action conservation limits

\backsection[Acknowledgements]{
I am thankful to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan
for the discussions that helped improve this paper.
}

@@

>>>>

\backsection[Acknowledgements]{
I am thankful to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan
for the discussions that helped improve this paper.
I also thank three anonymous reviewers for their helpful comments.
}

@@

\appendix
\section{Wave action conservation requirements}
\label{appendix:wave_action_conservation}

Following the variational approach by \citet{whitham1965general},
\citet{bretherton1968wavetrains} showed that the wave action is conserved for
small amplitude (linear) waves propagating in slowly-varying media.
Although \citet{longuet1987propagation} correctly pointed out both requirements
(that the medium is slowly-varying and that the short-wave properties are
approximately homogeneous and stationary) for the wave action conservation to
hold, he incorrectly stated that there is no explicit restriction on the
steepness of the long wave, and that it appears necessary to only assume
small-amplitude short waves and significant separation between $k$ and $k_L$.
We will show that in the general case, both the homogeneity and the stationarity
of a short wave quantity $q$ are dependent on the long-wave steepness
$\varepsilon_L$.

The homogeneity and stationarity of a scalar quantify $q$ can be expressed as
conditions on the instantaneous fractional rate of change of $q$ being much
smaller than the short-wave inverse scales:

\begin{equation}
  \frac{1}{q} \frac{\partial q}{\partial x} \ll k
  \label{eq:homogeneity_condition}
\end{equation}

\begin{equation}
  \frac{1}{q} \frac{\partial q}{\partial t} \ll \sigma
  \label{eq:stationarity_condition}
\end{equation}
The slowly-varying conditions on the medium (in our case, the long-wave orbital
velocity $U$) are then
\begin{equation}
  \frac{1}{U} \frac{\partial U}{\partial x} \ll \widetilde{k},
\end{equation}
and

\begin{equation}
  \frac{1}{U} \frac{\partial U}{\partial t} \ll \widetilde{\sigma},
\end{equation}
respectively.
For a linear long wave, these conditions are equivalent to the scale separation
between the long and short waves:

\begin{equation}
  k_L \ll \widetilde{k}
  \label{eq:k_scale_separation}
\end{equation}

\begin{equation}
  \sigma_L \ll \widetilde{\sigma}
\end{equation}
\citet{bretherton1968wavetrains} also require that the medium is nearly uniform
in the vertical, over the vertical scales of motion of the short waves.
For the ambient velocity being the orbital velocity of the long wave, this is
trivially satisfied with (\ref{eq:k_scale_separation}).

Now, to address the variation of short-wave properties,
we apply (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) to
the long-wave phase dependent wavenumber, gravitational acceleration, and
wave action, and define the homogeneity and stationarity quantities as:

\begin{equation}
  H_q = 1 - \left| \frac{1}{qk} \frac{\partial q}{\partial x} \right|
\end{equation}

\begin{equation}
  S_q = 1 - \left| \frac{1}{q\sigma} \frac{\partial q}{\partial t} \right|
\end{equation}
With the above definitions, we consider $q$ to be homogeneous and stationary as
$H_q \rightarrow 1$ and $S_q \rightarrow 1$, respectively.

For the linearized modulation solutions (\ref{eq:k_short_exact}),
(\ref{eq:gravity_modulation_linear_no_curvature_mean_level}), and
(\ref{eq:wave_action_modulation}), the respective homogeneity expressions are:

\begin{equation}
  H_{k,N} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right)^2} \right|
\end{equation}

\begin{equation}
  H_{g} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L^2 \cos^2{\psi}\right)} \right|
\end{equation}

\begin{equation}
  S_{k,N} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
\end{equation}

\begin{equation}
  S_{g} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
\end{equation}

\section{Surface velocity errors}
\label{appendix:surface_velocity_errors}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_phase.pdf}
\caption{
}
\label{fig:velocity_error_by_phase}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_relative_velocity_error.pdf}
\caption{
}
\label{fig:relative_velocity_error}
\end{figure}
